---
title: "Data Simulation"
output: html_notebook
---

# Simulating data for replication of Johnson & Newport (1989) 

The confirmatory hypothesis tests undertaken by Johnson & Newport (1989) were independent samples t-tests which compared the differences between two groups at a time.

The power analyses revealed that I will need a sample of n1 = 600 and n2 = 498 to replicate the original effect size with 80% power. The other sample sizes will not be changed.

## Groups means and sd

First, I will set the means and standard deviations of the sample distribution in accordance to those from the original study. Then, I will use the rnorm function to simulate mean test scores.

```{r}
native <- rnorm(545, mean = 268.8, sd = 2.9)
three_seven <- rnorm(545, mean = 269.3, sd = 2.8)
eight_ten <- rnorm(545, mean = 256, sd = 6)
eleven_fifteen <- rnorm(545, mean = 235.9, sd = 13.6)
seventeen_thirtynine <- rnorm(545, mean = 210.3, sd = 22.8)
```

## Setting the limits

Test scores have a lower limit of 0 and an upper limit of 276, so I will make sure none of the numbers exceed these restrictions.

```{r}
native[native > 276] <- 276
native[native < 0] <- 0
three_seven[three_seven > 276] <- 276
three_seven[three_seven < 0] <- 0
eight_ten[eight_ten > 276] <- 276
eight_ten[eight_ten < 0] <- 0
eleven_fifteen[eleven_fifteen > 276] <- 276
eleven_fifteen[eleven_fifteen < 0] <- 0
seventeen_thirtynine[seventeen_thirtynine > 276] <- 276
seventeen_thirtynine[seventeen_thirtynine < 0] <- 0
```

## Checking the means

```{r}
mean(native)
mean(three_seven)
mean(eight_ten)
mean(eleven_fifteen)
mean(seventeen_thirtynine)
```

## Creating the dataframe

I will use the gather function to horizontally stack data, and I will create a group factor and the score outcome variable.

```{r}
library(tidyverse)
data <- data.frame(ID=1:545, native=native, three_seven=three_seven, eight_ten=eight_ten, eleven_fifteen=eleven_fifteen, seventeen_thirtynine=seventeen_thirtynine)

data <- gather(data, key=group, value=score, native, three_seven, eight_ten, eleven_fifteen, seventeen_thirtynine)

data$group = as.factor(data$group)

data
```

## Save as dataframe (.csv)

Finally, I will save the dataframe as a .csv file

```{r}
write.csv(data, "data.csv")
```