---
title: "Confirmatory analysis"
output: html_notebook
---

## Replication study and hypotheses

This is a data analysis plan for the replication of:

Johnson, J. S., & Newport, E. L. (1989). Critical Period Effects in Second Language Learning: The Influence of Maturational State on the Acquisition of English as a Second Language. Cognitive Psychology, 21, 60–99. 


The original test of the primary hypothesis is a independent t-test that compares the mean grammar test score across five groups: native English speakers and second-language speakers, classified by the age at which they started to learn English. A power analysis revealed that 545 participants were needed in each condition to achieve at least 80% power. Here I will document the analysis plan for this replication. A successful replication will find evidence of support for the alternate hypothesis, which is:

H1: There will be a negative relationship between the age of exposure to English and the performance in the test, lasting up to adulthood.

The null hypothesis being tested in the independent t-tests is:

H0: There will be no relationship between the age of exposure to English and the performance in the test.

## Load data

```{r}
install.packages("vctrs")
library(vctrs)
install.packages('readr')
library(readr)
data <- read_csv("~/Desktop/45160/Data/data.csv")
```

## Preliminary analysis

### Data Screening and cleaning

First the data will be screened for erroneous and missing values. The outcome in this study, mean score, is bounded between 0 and 276 and so any responses beyond this range will be replaced with a missing value.

```{r}
library(tidyverse)
library(psych)
library(naniar)
data <-
  data %>%
  replace_with_na(replace = list(~.score > 276, ~.score < 0)) 
describe(data)
```

As this is an experimental study, any missing values will be removed listwise from the sample..

```{r}
drop_na(data) 
```

### Summary statistics and distributions

I will inspect the means and standard deviations of each group in the study.

```{r}
describeBy(data,data$group)
```

Then I will create a plot for the distributions of each group.

```{r}
library(ggplot2)
library(plyr)

## Find mean for each condition

means <- ddply(data, "group", summarise, score.mean=mean(score))

## Plot distributions

plot1 <- ggplot(data, aes(x=score)) + geom_histogram(binwidth=.5, colour="black", fill="white") + 
    facet_grid(group ~ .) +
    geom_vline(data=means, aes(xintercept=score.mean),
               linetype="dashed", size=1, colour="red") + theme_classic()

print(plot1)


## Save plot
ggsave(file="~/Desktop/45160/Plots/plot_1.png")
```

And finally create an APA table with the mean and sd for each condition.

```{r}
library(apaTables)
apa.1way.table(iv=group,dv=score,data=data,filename="~/Desktop/45160/Tables/table_1.doc",table.number = 1)
```

In order to perform 4 independent t-tests, I will use the split() function in order to split the 'data' data frame into 5 sub-data frames for each group. I will then combine the data frames two by two so that I can perform a t-test on each pair.

```{r}
list_of_dfs <- split(data, data$group)
df_native <- list_of_dfs$native
df_three_seven <- list_of_dfs$three_seven
df_eight_ten <- list_of_dfs$eight_ten
df_eleven_fifteen <- list_of_dfs$eleven_fifteen
df_seventeen_thirynine <- list_of_dfs$seventeen_thirtynine
df_native
df_three_seven
df_eight_ten
df_eleven_fifteen
df_seventeen_thirynine
```

The following code will combine the data frames two by two, producing 4 data frames ready to t-test.
```{r}
df_ttest1<- rbind(df_native, df_three_seven)
df_ttest2 <- rbind(df_three_seven, df_eight_ten)
df_ttest3 <- rbind(df_eight_ten, df_eleven_fifteen)
df_ttest4 <- rbind(df_eleven_fifteen, df_seventeen_thirynine)
df_ttest1
df_ttest2
df_ttest3
df_ttest4
```


## Confirmatory analysis

The next step is to conduct the independent samples t-test for the primary confirmatory hypothesis testing as listed in the data analysis plan. 

### T-test 1

I'll perform Levene's Test for Homogeneity of Variance to test the equality of variances between the groups.
```{r}
library(car)
levene_ttest1 <- leveneTest(score ~ group, data=df_ttest1)
levene_ttest1
```

The p-value is 0.737, which is greater than a 0.05 significance level. I cannot reject the hypothesis of equal variances, so I will use the var.equal argument set to TRUE in the t-test.

```{r}
library(MKinfer)
t.test(score ~ group, data=df_ttest1, var.equal=TRUE, paired=FALSE)
boot.t.test(score ~ group, data=df_ttest1, var.equal=TRUE, R = 5000)
```

Now I will calculate the effect size for the experimental condition.

```{r}
library(rstatix)
cohens_d(df_ttest1, score ~ group,  var.equal = TRUE)
```

#### Tables and figures for reporting 

To report the results from the independent t-test I will visualize the means and 95% confidence intervals for each condition using a bar chart with error bars.

```{r}
library(rcompanion)
sum <- groupwiseMean(score ~ group, data=df_ttest1,conf   = 0.95,digits = 3)
qplot(x    = group ,
      y    = Mean,
      data = sum) +

geom_bar(stat ="identity",
         fill = "gray", width = .40, alpha=.5
        ) +

geom_errorbar(aes(ymin  = Trad.lower,
                  ymax  = Trad.upper,
                  width = 0.15)) +
  xlab("Group") +
  ylab("Mean Test Score")+
  theme_classic()

ggsave(file="~/Desktop/45160/Plots/plot_2.png")
```
### T-test 2

I'll perform Levene's Test for Homogeneity of Variance to test the equality of variances between the groups.
```{r}
library(car)
levene_ttest2 <- leveneTest(score ~ group, data=df_ttest2)
levene_ttest2
```

The p-value is smaller than 2.2e-16, which is much lower than a 0.05 significance level. There is strong evidence to reject the hypothesis of equal variances, so I will use a Welch's t-test.

```{r}
library(MKinfer)
t.test(score ~ group, data=df_ttest2, var.equal=FALSE, paired=FALSE)
boot.t.test(score ~ group, data=df_ttest2, var.equal=FALSE, R = 5000)
```

Now I will calculate the effect size for the experimental condition.

```{r}
library(rstatix)
cohens_d(df_ttest2, score ~ group,  var.equal = FALSE)
```

#### Tables and figures for reporting 

To report the results from the independent t-test I will visualize the means and 95% confidence intervals for each condition using a bar chart with error bars.

```{r}
library(rcompanion)
sum <- groupwiseMean(score ~ group, data=df_ttest2,conf   = 0.95,digits = 3)
qplot(x    = group ,
      y    = Mean,
      data = sum) +

geom_bar(stat ="identity",
         fill = "gray", width = .40, alpha=.5
        ) +

geom_errorbar(aes(ymin  = Trad.lower,
                  ymax  = Trad.upper,
                  width = 0.15)) +
  xlab("Group") +
  ylab("Mean Test Score")+
  theme_classic()

ggsave(file="~/Desktop/45160/Plots/plot_3.png")
```

### T-test 3

I'll perform Levene's Test for Homogeneity of Variance to test the equality of variances between the groups.
```{r}
library(car)
levene_ttest3 <- leveneTest(score ~ group, data=df_ttest3)
levene_ttest3
```

The p-value is smaller than 2.2e-16, which is much lower than a 0.05 significance level. There is strong evidence to reject the hypothesis of equal variances, so I will use a Welch's t-test.

```{r}
library(MKinfer)
t.test(score ~ group, data=df_ttest3, var.equal=FALSE, paired=FALSE)
boot.t.test(score ~ group, data=df_ttest3, var.equal=FALSE, R = 5000)
```

Now I will calculate the effect size for the experimental condition.

```{r}
library(rstatix)
cohens_d(df_ttest3, score ~ group,  var.equal = FALSE)
```

#### Tables and figures for reporting 

To report the results from the independent t-test I will visualize the means and 95% confidence intervals for each condition using a bar chart with error bars.

```{r}
library(rcompanion)
sum <- groupwiseMean(score ~ group, data=df_ttest3,conf   = 0.95,digits = 3)
qplot(x    = group ,
      y    = Mean,
      data = sum) +

geom_bar(stat ="identity",
         fill = "gray", width = .40, alpha=.5
        ) +

geom_errorbar(aes(ymin  = Trad.lower,
                  ymax  = Trad.upper,
                  width = 0.15)) +
  xlab("Group") +
  ylab("Mean Test Score")+
  theme_classic()

ggsave(file="~/Desktop/45160/Plots/plot_4.png")
```

### T-test 4

I'll perform Levene's Test for Homogeneity of Variance to test the equality of variances between the groups.
```{r}
library(car)
levene_ttest4 <- leveneTest(score ~ group, data=df_ttest4)
levene_ttest4
```

The p-value is smaller than 2.2e-16, which is much lower than a 0.05 significance level. There is strong evidence to reject the hypothesis of equal variances, so I will use a Welch's t-test.

```{r}
library(MKinfer)
t.test(score ~ group, data=df_ttest4, var.equal=FALSE, paired=FALSE)
boot.t.test(score ~ group, data=df_ttest4, var.equal=FALSE, R = 5000)
```

Now I will calculate the effect size for the experimental condition.

```{r}
library(rstatix)
cohens_d(df_ttest4, score ~ group,  var.equal = FALSE)
```

#### Tables and figures for reporting 

To report the results from the independent t-test I will visualize the means and 95% confidence intervals for each condition using a bar chart with error bars.

```{r}
library(rcompanion)
sum <- groupwiseMean(score ~ group, data=df_ttest4,conf   = 0.95,digits = 3)
qplot(x    = group ,
      y    = Mean,
      data = sum) +

geom_bar(stat ="identity",
         fill = "gray", width = .40, alpha=.5
        ) +

geom_errorbar(aes(ymin  = Trad.lower,
                  ymax  = Trad.upper,
                  width = 0.15)) +
  xlab("Group") +
  ylab("Mean Test Score")+
  theme_classic()

ggsave(file="~/Desktop/45160/Plots/plot_5.png")
```
